Attrition in Online Learning: Understanding Persistence and Dropout in Massive Open Online Courses

### 2014-10-27 paper_v1.Rnw
added figure, made changes to methods and tables

% Load basic packages
\usepackage{balance}  % to better equalize the last page
\usepackage{graphics} % for EPS, load graphicx instead
\usepackage{times}    % comment if you want LaTeX's default font
\usepackage{url}      % llt: nicely formatted URLs
\usepackage{booktabs}

@@

TODO: Outline different courses and go through variabels in the course info table.

Table:  course name; enrollment; pre-survey response rate; prop. demographics (age, gender, educ); who intended to do all; post-survey response rate; self-identified dropout rate

@@

Table: proportions for each dropout reason for each course

>>>>

% Load basic packages
\usepackage{balance}  % to better equalize the last page
\usepackage{graphics} % for EPS, load graphicx instead
\usepackage{times}    % comment if you want LaTeX's default font
\usepackage{url}      % llt: nicely formatted URLs
\usepackage{booktabs}
\usepackage{rotating}

@@

% Which courses did we use? What data
Seventeen MOOCs on topcis in a variety of disciplines were selected for investigation: four on topics in the humanities, two in engineering, three in computer science, two runs of the same course in business, and six in mathematics and statistics. For each course, we merged data from the course survey (distributed in the first weeks of the course), the feedback surveys (distributed in the last weeks of the course), and behavioral data collected in the learning environment. Only those learners who completed the feedback survey were included in the analyses (some did not complete the initial course survey). Table \ref{tab:study1_courses_summary} provides the following summary statistics for each course: the number of enrolled learners, the number of learners who completed the course survey and the feedback surveys, age, gender, and education level, and the proportion of learners who self-reported that they stopped participating in the course. We also computed the proprtions of learners who viewed at least 30, 50, and 80\% of the course videos.

%Table:  course name; enrollment; pre-survey response rate; prop. demographics (age, gender, educ); who intended to do all; post-survey response rate; self-identified dropout rate

\begin{table*}
\label{tab:study1_courses_summary}
\caption{Summary statistics of 14 MOOCs}
\small
\center
\begin{tabular}{llcccccccccc}
\toprule
%  & Topic Area & Enrollment & Course survey & Feedback survey & Female & Age & Education & \specialcellcentered{\ensuremath{>} 30\% \\ videos} & \specialcellcentered{\ensuremath{>} 50\% \\ videos} & \specialcellcentered{\ensuremath{>} 80\% \\ videos} & \specialcellcentered{Self-Rep. \\ Dropout} \\
& & & & & & & & $>$ 30\% & $>$ 50\% & $>$ 80\% & Self-Rep. \\
& Topic Area & Enrollment & Course survey & Feedback survey & Female & Age & Education & videos & videos & videos & Dropout \\
\midrule
C1 & Business & 61,096 & 9,943 & 2,367 & 42\% & 39 (30,49) & n/a & 39\% & 20\% & 7\% & 10\% \\
C2 & Business & 41,081 & 6,544 & 883 & 40\% & 39 (30,49) & 97\% & 32\% & 13\% & 2\% & 9\% \\
C3 & CS & 27,104 & 6,104 & 2,854 & 35\% & 30 (22,41) & 74\% & 67\% & 40\% & 7\% & 2\% \\
C4 & CS & 56,154 & 3,406 & 1,123 & 11\% & 33 (27,42) & 91\% & 34\% & 16\% & 11\% & 21\% \\
C5 & CS & 57,636 & 5738 & 402 & 17\% & 33 (26,43) & 91\% & 66\% & 55\% & 28\% & 14\% \\
C6 & Engineering & 16,220 & 2,540 & 269 & 13\% & 33 (27,45) & 92\% & 62\% & 49\% & 17\% & 15\% \\
C7 & Enginerring & 7,178 & 1,312 & 202 & 12\% & 27 (22,39) & 85\% & 51\% & 18\% & 2\% & 7\% \\
C8 & Humanities & 7,634 & 3,631 & 152 & 92\% & 33 (26,45) & 93\% & 45\% & 11\% & 0\% & 7\% \\
C9 & Humanities & 6,475 & 3,315 & 326 & 90\% & 32 (27,47) & 91\% & 69\% & 50\% & 17\% & 16\% \\
C10 & Humanities & 9,903 & 4,227 & 1,570 & 52\% & 39 (29,52) & 89\% & 84\% & 74\% & 21\% & 3\% \\
C11 & Humanities & 630 & 167 & 80 & 50\% & 30 (27,40) & 97\% & 74\% & 55\% & 16\% & 16\% \\
C12 & Math \& Stats & 49,317 & 3,090 & 862 & 33\% & 41 (29,55) & 88\% & 27\% & 10\% & 1\% & 31\% \\
C13 & Math \& Stats & 32,006 & 26,142 & 5,336 & 58\% & 28 (16,46) & n/a & 69\% & 38\% & 8\% & 4\% \\
C14 & Math \& Stats & 16,857 & 2,892 & 363 & 9\% & 31 (26,38) & 99\% & 61\% & 45\% & 25\% & 14\% \\
C15 & Math \& Stats & 37,017 & 11,629 & 1,854 & 20\% & 35 (29,45) & 98\% & 78\% & 67\% & 34\% & 8\% \\
C16 & Math \& Stats & 27,236 & 2,711 & 637 & 82\% & 43 (34,51) & n/a & 64\% & 36\% & 5\% & 1\% \\
C17 & Math \& Stats & 13,132 & 3,629 & 843 & n/a & n/a & n/a & 70\% & 54\% & 14\% & 5\% \\
\bottomrule
\end{tabular}
\end{table*}

% What LMS data variables did we compute?
We are interested in exploring correlations between reasons for dropout and learner behaviors. If such correlations exist, they can be an additional towards understanding why learners engage with or disengage from various learning activities. In this study, we computed the fraction of videos watched by each learner, fraction of problems attempted, average grade scored on attempted problems, and number of forum posts and replies posted.

@@

%Table: proportions for each dropout reason for each course
<<prop-reason, echo=FALSE, fig.env='figure*', fig.pos='ht', fig.height=4, fig.width=9, fig.cap='Proportion of learners who indicated each reason for attrition across 17 MOOCs'>>=
values = c(18,20,10,22,19,21,8,11,13,13,9,7,14,18,11,14,19,
57,39,13,48,42,51,33,60,65,42,18,52,21,68,39,28,68,
21,23,25,14,13,19,8,0,0,23,9,12,20,20,13,28,12,
32,18,25,28,32,28,25,11,20,27,27,19,19,22,18,0,11,
15,14,8,16,13,12,0,0,9,10,9,9,19,12,7,14,14,
21,3,18,12,17,24,25,30,30,10,36,13,20,25,12,0,21,
69,58,38,76,80,66,66,55,69,67,27,71,28,64,54,0,71,
12,7,27,8,4,18,41,11,6,10,27,10,24,14,11,28,7,
18,7,26,17,19,16,41,10,9,25,18,21,27,25,20,28,12,
13,12,10,19,19,21,0,11,7,13,9,9,11,16,9,14,19,
16,5,22,17,14,21,16,11,7,18,20,16,23,20,16,16,12,
14,10,10,23,24,16,8,11,2,13,18,19,10,25,17,0,29,
26,8,11,27,23,39,33,11,24,11,27,27,11,35,15,14,41,
14,9,18,10,10,18,25,22,4,5,18,7,20,22,14,14,25)

reasons = c("Material not explained properly","Course requires too much time","Not what I was looking for","I joined too late","Didn't enjoy online course format","Technical difficulties","Couldn't keep up with deadlines","Not challenging enough","I learned all I wanted","Videos were confusing","I was only exploring","Too advanced for my background","Assignment or exam was too difficult","Credentials not valuable to me")

dat = data.frame(reasons = rep(reasons, each=17), values)

library(ggplot2)
ggplot(dat, aes(reorder(reasons, values, median), values)) + geom_boxplot() + coord_flip() + theme_bw() + labs(y="Proportion",x="")
@

% \begin{sidewaystable}
% \label{tab:proprtion_reason_for_dropout}
% \caption{Distribution of reported reasons for dropping out across courses. Percentage of learners who selected each reason for dropping out in each course, and median and interquartile range across courses}
% \small
% \center
% \begin{tabular}{llrrrrrrrrrrrrrrrrrr}
% \toprule
% Reason for dropout & C1 & C2 & C3 & C4 & C5 & C6 & C7 & C8 & C9 & C10 & C11 & C12 & C13 & C14 & C15 & C16 & C17 & Median & IQR \\
% \midrule
% \specialcellleft{Material not explained \\ properly} & 18 & 20 & 10 & 22 & 19 & 21 & 8 & 11 & 13 & 13 & 9 & 7 & 14 & 18 & 11 & 14 & 19 & 14 & 8 \\
% \specialcellleft{Course requires too \\ much time} & 57 & 39 & 13 & 48 & 42 & 51 & 33 & 60 & 65 & 42 & 18 & 52 & 21 & 68 & 39 & 28 & 68 & 42 & 24 \\
% \specialcellleft{Not what I was looking \\ for} & 21 & 23 & 25 & 14 & 13 & 19 & 8 & 0 & 0 & 23 & 9 & 12 & 20 & 20 & 13 & 28 & 12 & 14 & 9 \\
% I joined too late & 32 & 18 & 25 & 28 & 32 & 28 & 25 & 11 & 20 & 27 & 27 & 19 & 19 & 22 & 18 & 0 & 11 & 22 & 9 \\
% \specialcellleft{Didn't enjoy online \\ course format} & 15 & 14 & 8 & 16 & 13 & 12 & 0 & 0 & 9 & 10 & 9 & 9 & 19 & 12 & 7 & 14 & 14 & 12 & 5 \\
% Technical difficulties & 21 & 3 & 18 & 12 & 17 & 24 & 25 & 30 & 30 & 10 & 36 & 13 & 20 & 25 & 12 & 0 & 21 & 20 & 13 \\
% \specialcellleft{Couldn't keep up with \\ deadlines} & 69 & 58 & 38 & 76 & 80 & 66 & 66 & 55 & 69 & 67 & 27 & 71 & 28 & 64 & 54 & 0 & 71 & 66 & 15 \\
% Not challenging enough & 12 & 7 & 27 & 8 & 4 & 18 & 41 & 11 & 6 & 10 & 27 & 10 & 24 & 14 & 11 & 28 & 7 & 11 & 16 \\
% I learned all I wanted & 18 & 7 & 26 & 17 & 19 & 16 & 41 & 10 & 9 & 25 & 18 & 21 & 27 & 25 & 20 & 28 & 12 & 19 & 9 \\
% Videos were confusing & 13 & 12 & 10 & 19 & 19 & 21 & 0 & 11 & 7 & 13 & 9 & 9 & 11 & 16 & 9 & 14 & 19 & 12 & 7 \\
% I was only exploring & 16 & 5 & 22 & 17 & 14 & 21 & 16 & 11 & 7 & 18 & 20 & 16 & 23 & 20 & 16 & 16 & 12 & 16 & 6 \\
% \specialcellleft{Too advanced for my \\ background} & 14 & 10 & 10 & 23 & 24 & 16 & 8 & 11 & 2 & 13 & 18 & 19 & 10 & 25 & 17 & 0 & 29 & 14 & 9 \\
% \specialcellleft{Assn or exam was too \\ difficult} & 26 & 8 & 11 & 27 & 23 & 39 & 33 & 11 & 24 & 11 & 27 & 27 & 11 & 35 & 15 & 14 & 41 & 24 & 16 \\
% \specialcellleft{Credentials not valuable \\ to me} & 14 & 9 & 18 & 10 & 10 & 18 & 25 & 22 & 4 & 5 & 18 & 7 & 20 & 22 & 14 & 14 & 25 & 14 & 10 \\
% \bottomrule
% \end{tabular}
% \end{sidewaystable}

### 2014-10-27 paper_v1.tex
added figure, made changes to methods and tables

Table: proportions for each dropout reason for each course

>>>>

%Table: proportions for each dropout reason for each course
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure*}[ht]

\includegraphics[width=\maxwidth]{figure/prop-reason} \caption[Proportion of learners who indicated each reason for attrition across 17 MOOCs]{Proportion of learners who indicated each reason for attrition across 17 MOOCs\label{fig:prop-reason}}
\end{figure*}

\end{knitrout}

% \begin{sidewaystable}
% \label{tab:proprtion_reason_for_dropout}
% \caption{Distribution of reported reasons for dropping out across courses. Percentage of learners who selected each reason for dropping out in each course, and median and interquartile range across courses}
% \small
% \center
% \begin{tabular}{llrrrrrrrrrrrrrrrrrr}
% \toprule
% Reason for dropout & C1 & C2 & C3 & C4 & C5 & C6 & C7 & C8 & C9 & C10 & C11 & C12 & C13 & C14 & C15 & C16 & C17 & Median & IQR \\
% \midrule
% \specialcellleft{Material not explained \\ properly} & 18 & 20 & 10 & 22 & 19 & 21 & 8 & 11 & 13 & 13 & 9 & 7 & 14 & 18 & 11 & 14 & 19 & 14 & 8 \\
% \specialcellleft{Course requires too \\ much time} & 57 & 39 & 13 & 48 & 42 & 51 & 33 & 60 & 65 & 42 & 18 & 52 & 21 & 68 & 39 & 28 & 68 & 42 & 24 \\
% \specialcellleft{Not what I was looking \\ for} & 21 & 23 & 25 & 14 & 13 & 19 & 8 & 0 & 0 & 23 & 9 & 12 & 20 & 20 & 13 & 28 & 12 & 14 & 9 \\
% I joined too late & 32 & 18 & 25 & 28 & 32 & 28 & 25 & 11 & 20 & 27 & 27 & 19 & 19 & 22 & 18 & 0 & 11 & 22 & 9 \\
% \specialcellleft{Didn't enjoy online \\ course format} & 15 & 14 & 8 & 16 & 13 & 12 & 0 & 0 & 9 & 10 & 9 & 9 & 19 & 12 & 7 & 14 & 14 & 12 & 5 \\
% Technical difficulties & 21 & 3 & 18 & 12 & 17 & 24 & 25 & 30 & 30 & 10 & 36 & 13 & 20 & 25 & 12 & 0 & 21 & 20 & 13 \\
% \specialcellleft{Couldn't keep up with \\ deadlines} & 69 & 58 & 38 & 76 & 80 & 66 & 66 & 55 & 69 & 67 & 27 & 71 & 28 & 64 & 54 & 0 & 71 & 66 & 15 \\
% Not challenging enough & 12 & 7 & 27 & 8 & 4 & 18 & 41 & 11 & 6 & 10 & 27 & 10 & 24 & 14 & 11 & 28 & 7 & 11 & 16 \\
% I learned all I wanted & 18 & 7 & 26 & 17 & 19 & 16 & 41 & 10 & 9 & 25 & 18 & 21 & 27 & 25 & 20 & 28 & 12 & 19 & 9 \\
% Videos were confusing & 13 & 12 & 10 & 19 & 19 & 21 & 0 & 11 & 7 & 13 & 9 & 9 & 11 & 16 & 9 & 14 & 19 & 12 & 7 \\
% I was only exploring & 16 & 5 & 22 & 17 & 14 & 21 & 16 & 11 & 7 & 18 & 20 & 16 & 23 & 20 & 16 & 16 & 12 & 16 & 6 \\
% \specialcellleft{Too advanced for my \\ background} & 14 & 10 & 10 & 23 & 24 & 16 & 8 & 11 & 2 & 13 & 18 & 19 & 10 & 25 & 17 & 0 & 29 & 14 & 9 \\
% \specialcellleft{Assn or exam was too \\ difficult} & 26 & 8 & 11 & 27 & 23 & 39 & 33 & 11 & 24 & 11 & 27 & 27 & 11 & 35 & 15 & 14 & 41 & 24 & 16 \\
% \specialcellleft{Credentials not valuable \\ to me} & 14 & 9 & 18 & 10 & 10 & 18 & 25 & 22 & 4 & 5 & 18 & 7 & 20 & 22 & 14 & 14 & 25 & 14 & 10 \\
% \bottomrule
% \end{tabular}
% \end{sidewaystable}
